Impact: An Indirect Dependency Awareness Tool For Software Development

### 2013-01-19 icse-chase-2013/paper.tex
Danas revision 3 is now final.

% Paper abstract
\begin{abstract}
Awareness has been largely studied in the field of awareness research within
software engineering. Many tools and techniques been been proposed and built
in order to provide software developers and stakeholders with a greater
sense of workspace and task awareness within their software projects. These
techniques and tools have been largely focused on detecting \textit{direct} 
conflicts which arise over a project's life time or have created 
an exploratory ground for stakeholders to use as a means of resolving self
discovered direct or \textit{indirect} conflicts. However, detecting 
and providing pertinent information regarding indirect conflicts has been
largely ignored partially due to its inherently larger
complexity than direct conflicts. Indirect conflicts arise when changes
in one software artifact affect another. In this paper, we present
\textit{Impact}, a new task awareness tool directly aimed at both detecting
and presenting indirect conflicts which arise inside of a software project.
We introduce previous indirect conflict awareness attempts, our design 
and implementation, as well as describe \textit{Impact's} potential through
two evaluation studies. 
\end{abstract}

@@

Despite software developer's need for indirect conflict awareness tools
detecting and alerting developers to arising indirect conflicts is still a major
problem. \textit{Impact}, a web based awareness
tool, aims to solve this issue. In this paper, we describe \textit{Impact's}
design and implementation in order to both detect newly created
indirect conflicts among software developers as well as alerting developers
to these conflicts. By leveraging technical relationships inherent of 
software projects with method call graphs~\cite{Lakhotia:1993:CCM}
as well as detecting changes
to these technical relationship through software configuration management
(SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as
alert developers involved in such conflicts in task awareness. While this paper
outlines \textit{Impact's} specific implementation, the design used can be put
towards many other similar indirect conflict awareness tools. This design
fulfills the technical and social dependency congruence found by 
Grinter  et al.~\cite{Grinter:2003:RCW} which suggests to offer a better 
communication patterns inside the project.

@@

We asked lead developers to address two main concerns: do indirect
conflicts pose a threat at the method level, and did \textit{Impact} help raise
awareness and promote quicker conflict resolution for indirect
conflicts. Our two interviews largely supported our initial claims of
indirect conflicts posing a serious threat to developers, especially
in medium to large teams or projects as opposed to the small
teams which they were apart of. Our interviews also pointed
out that method use can be a particularly large area for indirect
conflicts to arise. However, both interviewees also pointed out that
any technical object which is used as an interface to some data
construct or methodology, database access for instance, can be 
a large potential issue for indirect conflicts.  Interview response to
\textit{Impact} was also largely positive, as both interviewees said that \textit{Impact}
helped raise awareness among their teams with what other developers
are doing as well as the influence it has on their own work. However,
both interviews showed \textit{Impact} to have information overload. Both
interviews showed that while all method changes were being detected,
not all are alert worthy. One lead developer suggested to only alert
developers to indirect conflicts if the internal structure of a method
changes due to modification to input parameters or output parameters.
In other words, the boundaries of the technical objects (changing
how a parameter is used inside the method, modifying the return
result inside the method) seem to be more of interest than other 
internal workings.

These two studies have shown that our design and approach to
detecting and alerting developers to indirect conflicts appear
to be on the correct path. \textit{Impact} as a tool has laid the foundations
for future work in detecting indirect conflicts as well as alerting
developers, although more thought must be given as to 
what constitutes a meaningful change inside our selected 
technical objects.

\section{Conclusion and Future Work}
In this paper, we have presented the issues that arise from indirect 
conflicts in present awareness tools. We have proposed a generic 
design for the future development of awareness tools in regards to
handling indirect conflicts. We have presented a prototype 
awareness tool, \textit{Impact}, which was designed around our generic 
awareness approach. \textit{Impact} was evaluated on a small scale, showing
its future potential as well as highlighting its current weaknesses.

In future work we are planning to conduct interviews and surveys
with software developers to confirm that indirect conflicts pose a
threat to their projects as well as discovering what constitutes a 
valid change inside of a given technical object. With these two
improvements, we plan to justify our generic design of indirect
conflict tools further as well as improve the ideas of detecting 
valid and significant changes to technical objects.

>>>>

% Paper abstract
\begin{abstract}
Awareness has been largely studied in the field of awareness research within
software engineering. Many tools and techniques been been proposed and built
in order to provide software developers and stakeholders with a greater
sense of workspace and task awareness within their software projects. These
techniques and tools have been largely focused on detecting \textit{direct} 
conflicts which arise over a project's life time or have created 
an exploratory ground for stakeholders to use as a means of resolving self
discovered direct or \textit{indirect} conflicts. However, detecting 
and providing pertinent information regarding indirect conflicts has been
largely ignored partially due to its inherently larger
complexity than direct conflicts. Indirect conflicts arise when changes
in one software artifact affect another. In this paper, we present
\textit{Impact}, a new task awareness tool directly aimed at both detecting
and presenting indirect conflicts which arise inside of a software project.
\textit{Impact} represents a first step towards the design and implementation
of awareness tools that specifically address indirect conflicts in software
projects. To describe \textit{Impact}, we introduce previous indirect 
conflict awareness attempts, our design 
and implementation, as well as \textit{Impact's} potential through
two evaluation studies. 
\end{abstract}

@@

Despite software developer's need for indirect conflict awareness tools
detecting and alerting developers to arising indirect conflicts is still a major
problem. \textit{Impact}, a web based awareness
tool, aims to solve this issue. In this paper, we describe \textit{Impact's}
design and implementation in order to both detect newly created
indirect conflicts among software developers as well as alerting developers
to these conflicts. By leveraging technical relationships inherent of 
software projects with method call graphs~\cite{Lakhotia:1993:CCM}
as well as detecting changes
to these technical relationship through software configuration management
(SCM) systems, \textit{Impact} is able to detect indirect conflicts as well as
alert developers involved in such conflicts in task awareness. While this paper
outlines \textit{Impact's} specific implementation, the design used can be put
towards many other similar indirect conflict awareness tools. 
\textit{Impact} represents a first step towards the design and implementation
of awareness tools which address indirect conflicts in software development.

@@

We asked lead developers to address two main concerns: do indirect
conflicts pose a threat at the method level, and did \textit{Impact} help raise
awareness and promote quicker conflict resolution for indirect
conflicts. Our two interviews largely supported our initial claims of
indirect conflicts posing a serious threat to developers, especially
in medium to large teams or projects as opposed to the small
teams which they were apart of. It was also pointed
out that method use can be a particularly large area for indirect
conflicts to arise. However, it was noted that
any technical object which is used as an interface to some data
construct or methodology, database access for instance, can be 
a large potential issue for indirect conflicts.  Interview response to
\textit{Impact} was also largely positive, as interviewees stated that \textit{Impact}
helped raise awareness among their teams with what other developers
are doing as well as the influence it has on their own work. However,
\textit{Impact} was shown to have information overload. It was suggested
that while all method changes were being detected,
not all are notification worthy. One developer suggested to only notify
developers to indirect conflicts if the internal structure of a method
changes due to modification to input parameters or output parameters.
In other words, the boundaries of the technical objects (changing
how a parameter is used inside the method, modifying the return
result inside the method) seem to be more of interest than other 
internal workings. More complex inner workings of methods were also noted 
to be of interest to developers
such as cyclomatic complexity, or time and space requirements.

These two studies have shown that our design and approach to
detecting and alerting developers to indirect conflicts appear
to be on the correct path. \textit{Impact} as a tool has laid the foundations
for future work in detecting indirect conflicts as well as notifying
developers, although more thought must be given as to 
what constitutes a meaningful change inside our selected 
technical objects.

\section{Conclusion and Future Work}
In this paper, we have presented the issues that arise from indirect 
conflicts in present awareness tools. We have proposed a generic 
design for the future development of awareness tools in regards to
handling indirect conflicts. We have presented a prototype 
awareness tool, \textit{Impact}, which was designed around our generic 
awareness approach. \textit{Impact} was evaluated on a small scale, showing
its future potential as well as highlighting its current weaknesses. This paper
has demonstrated our first step towards understanding awareness tools
that specifically address indirect conflicts and what questions are 
still to be answered.

\section{Future Work For Indirect Conflicts}
While \textit{Impact} has laid the foundations for awareness tool construction
aimed at resolving indirect conflicts, it has also raised questions 
in regards to levels of granularity in notification of indirect conflicts and the
implied trade off between notifications and information overload. To address
these problems, we hope to find industry expert opinion through interviews
and surveys with experienced developers. From these interactions, we hope to
answer some of the questions presented in this paper such as which technical
objects concern developers with indirect conflicts, and what types of changes
to these objects are cause for notification. We also hope, through these
expert opinions, to further the 
push for indirect conflict awareness tools by establishing that indirect
conflicts are real world problems with harmful consequences. 
Further more, we hope to bring these opinions back to the realm of 
\textit{Impact}, in combination with other academic works on balancing
between notifications and information overload~\cite{Wang:2007}, to
continue our development
on this tool to further its life both as a research tool and a possible
industry application.
